Choose a License

### 2024-09-13 choose_license.qmd
Add a TL;DR to the chapter on licenses

@@

## Which License to Choose?

@@

In most cases, the output of software, like images or tables, does not depend on its license. Therefore, if you use an R package under a copyleft license to create content, you are likely the copyright owner. However, if the output is based in data, it can be considered a derivative work and the license of the data also applies.

@@

@@

Having selected the licenses of your choice -- you might need multiple ones depending on the types of works your project contains --, we encourage you to read through the full license text (or at least a legal summary) to understand their effect. Then, you can record the license of existing content and apply the licenses of your own contributions. Mostly, this just means indicating which license applies to which file or folder, usually in the project's README (among other places), whose creation will be discussed in "[Add a README](add_readme.qmd)".

>>>>

## TL;DR

Every expression of an idea, such as literary or artistic works, are automatically covered by copyright. This means that nobody else than the copyright holder is allowed to copy, modify, or share it. Copyright licenses can grant others some of the necessary rights, but mostly don't cover publicity, privacy, moral, patent, or trademark rights.

Researchers are only allowed to share material they have the necessary rights to, for example, through a free/open license. These licenses often require to attribute the authors, indicate whether changes were made, and provide the text of the license, among other things. In addition, researchers should share any work by their own under such a license.

"Choosing a License" involves:

1. Recording the copyright notices and licenses for files by others (for example, in a file called `LICENSE.txt`) and following all their additional requirements.
2. In the simplest case, choosing two licenses (one for code, one for everything else) which to apply  simultaneously (at the choice of the recipient) to all own files (for example, again by recording the license in the file `LICENSE.txt`).

@@

## Which License to Choose for a Work?

@@

In most cases, the output of software, like images or tables, does not depend on its license. Therefore, if you use an R package under a copyleft license to create a figure, you are likely the copyright owner. However, if the output is based on data, it can be considered a derivative work and the license of the data also applies. For example, maps may be considered as a derivative of the geographic data they are based on.

@@

## Which Licenses to Choose for a Project?

So far, we only discussed how to choose a license for works _of one type_. But what if you want to share a project with all kinds of files? For example, the project from this tutorial (among other things) includes a data file, a manuscript file with intermingled code and writing, and an R file. And what if you also want to share files by others, as is the case with `apaquarto` which you may have installed in this project?

The answer is that you need to indicate the license on a per-file or per-folder basis (rather than choosing one for the whole project). The easiest approach is to make note of every foreign work included in your project and record its license. Then, dual-license all the remaining files, which are yours, under a code and a non-code license in parallel as explained in @tip-license-r.

@@

Having selected the licenses of your choice -- again, you might need multiple ones depending on the types of works your project contains --, we encourage you to read through the full license text (or at least a legal summary) to understand their effect. Then, you can record the license of existing content and apply the licenses of your own contributions. Mostly, this just means indicating which license applies to which file or folder, usually in the project's README (among other places), whose creation will be discussed "[later](make_readme.qmd)".
